### 2022-12-08 intro.Rnw
some suggestions Leo

% How can we know if a finding from a scientific study is really true? For
% example, how can we know if the protective effect of a vaccine found in a study
% is real? The answer to such questions is of high importance to scientists,
% decision makers, and the general population. However, because any study result
% comes with uncertainty we can never know for sure.
How can we know whether we can trust a finding from a scientific study? For
example, how can we know whether the protective effect of a vaccine found in a
study is a genuine effect or just a spurious finding? The answer to such
questions is of high importance to scientists, decision makers, and the general
population. However, because every study result is subject to uncertainty, we
can never know for sure.

One way to come a little closer to the truth is to repeat the original study
with new subjects. Such a \emph{replication study} may yield similar results
which would make us more confident about the original finding, or may yield
conflicting results which would lower our confidence. As such, replication
studies are an essential part of the scientific process as they provide a means
for substantiating genuine research findings and refuting research findings
which occurred merely by chance. For this reason, ``successful'' replication is
often a requirement, for instance, for acceptance of newly proposed scientific
theories (\eg{} a new physical model) or the implementation of policies based on
scientific knowledge (\eg{} market approval of a drug). The results from
replication studies may thus have real world consequences, such as deciding
whether we receive the recommendation to get vaccinated (a consequence that
anyone who experienced the COVID-19 pandemic is very aware of).

@@

In the midst of this crisis, % , particularly from the hard-hit
% field of psychology,
various reforms were implemented to prevent an escalation of the situation % and
% save the reputation of science
\citep[for an overview see \eg{}][]{Munafo2017}. For instance, many journals
adopted the ``registered report'' format \citep{Chambers2021} in which a study
proposal is peer reviewed \emph{before} the study is conducted, and which, if
accepted, gives provisional publication acceptance regardless of the study
outcome. Similarly, digital infrastructure platforms, such as Zenodo
(\url{https://zenodo.org}) or the Open Science Framework (\url{https://osf.io}),
were created to facilitate preregistration, preprints, code and data sharing,
all of which have substantially increased over the last decade
\citep{Kidwell2016, Nosek2018, Rawlinson2019}. The practice of conducting
replication studies has also gained popularity. Several academic journals and
funders are now explicitly promoting replication research \citep{NWO2016,
  NSL2018, Nature2022}. Figure~\ref{fig:scholar} (left) illustrates this trend
by the yearly number of results for the search term ``replication study''
obtained from the search engine Web of Science. The numbers have been rapidly
growing, especially since the mid-2000s. This trend could possibly be explained
by the fact that more research results are published each year but also when the
numbers are normalized by the number of results for the search term ``study'',
the increasing trend is still visible.

@@

\begin{table}[!htb]
  \centering
  \caption{Study-level summary statistics for an original and a replication
    study. The cumulative distribution function of the standard normal
    distribution is denoted by $\Phi(\cdot)$, and the $1 - \alpha$ quantile of
    the standard normal distribution is denoted by
    $\Phi^{-1}(1 - \alpha) = \z{\alpha}$. Confidence intervals and
    \textit{p}-values are based on the assumption that the effect estimate
    $\that$ is normally distributed around the unknown effect size $\theta$ with
    (known) variance equal to the squared standard error $\sigma^{2}$.}
  \label{tab:meta}
  \begin{tabular}{lcc}
    \toprule
    & Original study & Replication study \\
    \midrule
    effect estimate & $\that_o$ & $\that_{r}$ \\
    standard error & $\sigma_{o}$ & $\sigma_{r}$ \\
    $(1 - \alpha)$ confidence interval &
    $[\that_{o} \pm \z{\alpha/2} \sigma_{o}]$ &
    $[\that_{r} \pm \z{\alpha/2} \sigma_{r}]$ \\
    % $[\that_{o} - \z{\alpha/2} \sigma_{o}, \that_{o} + \z{\alpha/2} \sigma_{o}]$ &
    % $[\that_{r} - \z{\alpha/2} \sigma_{r}, \that_{r} + \z{\alpha/2} \sigma_{r}]$ \\
    $z$-value & $z_{o} = \that_{o}/\sigma_{o}$ & $z_{r} = \that_{r}/\sigma_{r}$ \\
    $p$-value (two-sided) & $p_{o} = 2\{1 - \Phi(|z_{o}|)\}$ & $p_{r} = 2\{1 - \Phi(|z_{r}|)\}$ \\
    % One-sided $p$-value & $p_{o} = 1 - \Phi(|z_{o}|)$ & $p_{r} = 2\{1 - \Phi(|z_{r}|)\}$ \\
    \bottomrule
  \end{tabular}
\end{table}

Most methods for analyzing replication studies can be formulated in terms of
study-level summary statistics as shown in Table~\ref{tab:meta}. % Effect
% estimates, standard error, confidence intervals, and $p$-values
All of these are routinely reported in research articles, and if one of them is
missing they typically can be calculated from the others. Using summary
statistics is also often the only possible way for conducting the replication
success analysis as the raw data from the original study may not be available to
the researchers conducting the replication study. The most important statistic
is the effect estimate $\that$. It provides an estimate of the underlying effect
size $\theta$ which quantifies the true effect or association of an
intervention/exposure with an outcome variable. Typical effect sizes are mean
differences and correlations (for continuous outcomes), odds ratios (for binary
outcomes), or hazard ratios (for time to event outcomes). Depending on the type
of effect size, a transformation might be required so that the assumption of
approximately normally distributed effect estimates around the unknown effect
size $\theta$ (for large enough sample sizes) is justified. This could be, for
instance, the Fisher $z$-transformation for correlations or the
log-transformation for odds/hazard ratios \citep[chapter 11]{Cooper2019}. The
associated standard error $\sigma$ represents the statistical uncertainty of the
estimate.\footnote{Typically, the available standard error $\sigma$ is only an
  estimate of the true standard error. However, in most parts of this thesis it
  will be assumed that $\sigma$ corresponds to the true standard error. This is
  the same assumption as in ordinary fixed-effect/random-effects meta-analysis,
  and typically reasonable if the sample size of the studies is not too small.}
Under the assumption of (asymptotic) normality, confidence intervals for
$\theta$ and $p$-values for testing the null hypothesis
$H_{0} \colon \theta = 0$ can be computed as shown in Table~\ref{tab:meta}.

@@

Table~\ref{tab:rsdef} lists commonly used criteria for replication success in
terms of the summary statistics from Table~\ref{tab:meta}. The most popular
criterion is listed first and defines replication success by simultaneous
statistical significance of original and replication study along with their
effect estimates showing the same direction. This approach is also called the
``two-trials rule'' \citep{Senn2008} in drug regulation or ``vote-counting'' in
meta-analysis \citep{Cooper2019}. The criterion can similarly be defined through
one-sided $p$-values, so that the original and replication effect estimate are
not required to show the same direction as this is taken into account by the
one-sided $p$-values. Some replication projects \citep{Opensc2015,
  Errington2021} also defined replication success via simultaneous
non-significance of original and replication study ($p_{o} > \alpha$ and
$p_{r} > \alpha$). However, with this definition ``replication success'' can
almost always be achieved by conducting original and replication study with very
few samples so that the $p$-values are large. The approach is also logically
questionable as it could be seen as an instance of the ``absence of evidence
fallacy'' \citep{Altman1995} meaning the failure to find evidence against the
null hypothesis is erroneously interpreted as evidence for the null hypothesis.
Meta-analytic extensions of the significance approach define replication success
by significance of a combined effect estimate.\footnote{The meta-analytic
  approach can also be given a Bayesian interpretation: A flat prior
  distribution for the underlying effect size $\theta$ is updated by the data
  from original and replication study. Replication success defined by a
  meta-analytic $p$-value being smaller than $\alpha$ is then equivalent to
  replication success via a Bayesian posterior probability
  $\Pr(\theta > 0 \given \that_{o}, \that_{r}) > 1 - \alpha/2$, respectively,
  $\Pr(\theta < 0 \given \that_{o}, \that_{r}) > 1 - \alpha/2$, depending on the
  direction of the combined estimate.} Typically, the assumption of a common
underlying effect size is seen as reasonable so that fixed-effect meta-analysis
is used for pooling. Random-effects meta-analysis has primarily been used if
more than one replication study of the same original study are conducted since
replicators are often interested in understanding between-replication
heterogeneity \citep[\eg{} in][]{Klein2018}.

@@

\begin{figure}[!htb]
<< "RPCB-graph", fig.height = 3.5 >>=

## function to format CI
ciFormat. <- function(lower, upper) {
    paste0("[", format(signif(lower, 2), nsmall = 2),
           ", ", format(signif(upper, 2), nsmall = 2), "]")
}
ciFormat <- Vectorize(FUN = ciFormat.)

## function to format p-value
pFormat. <- function(p, digits = 2, dollar = FALSE, phantom = FALSE) {
    if (is.na(p)) {
        out <- NA
    } else {
        if (p > 0.0001) {
            out <- as.character(signif(p, digits = digits))
            if (phantom) out <- paste0("\\phantom{<~}", out)
        } else {
            out <- "< 0.0001"
        }
        if (dollar) {
            out <- paste0("$", out, "$")
        }
    }
    return(out)
}
pFormat <- Vectorize(FUN = pFormat.)

za <- qnorm(p = 0.975)
rpcb <- read.csv("data/rpcb.csv") |>
    mutate(zo = smdo/so,
           zr = smdr/sr,
           c = so^2/sr^2,
           d = smdr/smdo,
           po2 = 2*(1 - pnorm(q = abs(zo))),
           pr2 = 2*(1 - pnorm(q = abs(zr))),
           sm = 1/sqrt(1/so^2 + 1/sr^2),
           smdm = (smdo/so^2 + smdr/sr^2)*sm^2,
           pm2 = 2*(1 - pnorm(q = abs(smdm/sm))),
           Q = (smdo - smdr)^2/(so^2 + sr^2),
           pQ = pchisq(q = Q, df = 1, lower.tail = FALSE),
           dFormat = tableFormat(x = d, outFun = dFormat,
                                 threshold = 1, col1 = "dark2orange", col2 = "dark2green"),
           cioL = smdo - za*so,
           cioU = smdo + za*so,
           cirL = smdr - za*sr,
           cirU = smdr + za*sr,
           cimL = smdm - za*sm,
           cimU = smdm + za*sm,
           id = paste0("(", paper, ", ", experiment, ", ", effect, ")"),
           po2Format = tableFormat(x = po2),
           pr2Format = tableFormat(x = pr2),
           pm2Format = tableFormat(x = pm2),
           pQFormat = tableFormat(x = pQ),
           smdoFormat = paste("$", dFormat(smdo, phantom = TRUE), "~",
                              ciFormat(lower = cioL, upper = cioU), "$"),
           smdrFormat = paste("$", dFormat(smdr, phantom = TRUE), "~",
                              ciFormat(lower = cirL, upper = cirU), "$"),
           smdmFormat = paste("$", dFormat(smdm, phantom = TRUE), "~",
                              ciFormat(lower = cimL, upper = cimU), "$"),
           rsig = factor(pr2 <= 0.05, levels = c(FALSE, TRUE),
                         labels = c("italic(p[r]) > 0.05",
                                    "italic(p[r]) <= 0.05")))
## compute sceptical p-value (TODO fix problems with NA)
ind <- !is.na(rpcb$zo) & !is.na(rpcb$zr) & !is.na(rpcb$c)
rpcb$ps2 <- NA
rpcb$ps2[ind] <- pSceptical(zo = rpcb$zo[ind], zr = rpcb$zr[ind], c = rpcb$c[ind],
                            alternative = "two.sided", type = "nominal")
rpcb$ps2Format <- NA
rpcb$ps2Format <- tableFormat(x = rpcb$ps2)

@@

upset(rpcbUpset, sets = c("Significance", "Meta analytic \nsignificance",
                          "Relative \neffect size", "Prediction \ninterval"),
      order.by = "freq", mainbar.y.label = "Successful replications (combined criteria)",
      sets.x.label = "Successful replications \n(single criterion)", #, empty.intersections = "on")
      queries = list(list(query = intersects,
                          params = c("Significance",
                                     "Meta analytic \nsignificance",
                                     "Relative \neffect size",
                                     "Prediction \ninterval"),
                          color = "#1b9e77", active = TRUE)))
@
\caption{UpSet plot for data on 132 effects from the Reproducibility Project:
  Cancer Biology \citep{Errington2021} for which effect estimates and standard
  errors are available on standardized mean difference scale. Shown are the
  number of replication successes according to the different criteria from
  Table~\ref{tab:rsdef} and their combinations. A threshold of
  $\alpha = \Sexpr{round(100*alpha, 2)}\%$ and a relative effect size threshold
  of $d_{\text{min}} = \Sexpr{dmin}$ are used for replication success.}
\label{fig:upset}
\end{figure}

@@

Taken together, this analysis demonstrates that conclusions based on commonly
used replication success criteria often differ. It is not always clear cut
whether or not a replication is successful. Reducing replicability to a single
criterion without mentioning these difficulties, as often done by the popular
press \citep[\eg{} ``more than half of the findings did not hold up when
retested'' in][]{Carey2015}, is a simplification of the matter.

\begin{landscape}
\begin{table}
  \centering
  \caption{Subset of results from the Reproducibility Project: Cancer Biology
    \citep{Errington2021}. (P, X, E) denotes effect number E from experiment
    number X, from original paper number P. Shown are original $\that_{o}$,
    replication $\that_{r}$, and pooled effect estimate $\that_{m}$ with 95\%
    confidence intervals, variance ratio $c = \sigma^{2}_{o}/\sigma^{2}_{r}$,
    relative effect estimate $d = \that_{r}/\that_{o}$, original
    \textit{p}-value $p_{o}$, replication \textit{p}-value $p_{r}$,
    meta-analytic \textit{p}-value $p_{m}$, $Q$-test \textit{p}-value $p_{Q}$,
    and sceptical \textit{p}-value $p_{\mathrm{S}}$. A replication success
    (green) threshold of $\alpha = 5\%$ is used for the \textit{p}-values and
    $d_{\text{min}} = 1$ for the relative effect estimate.}
  \label{tab:RPCB}
\resizebox{1\linewidth}{!}{%
<< "RPCB-table2", results = "asis" >>=
## select a subset with one effect per paper
set.seed(7)
rpcbTable <- rpcb |>
    na.omit() |>
    ## filter(smdo < 80, smdr < 80) |>
    group_by(paper) |>
    sample_n(size = 1) |>
    ungroup() |>
    arrange(pr2) |>
    select(id,
           smdoFormat, ## so,
           smdrFormat, smdmFormat,
           c, dFormat, ## sr,
           po2Format, pr2Format, pm2Format, pQFormat, ps2Format
           )

rpcbXtable <- xtable(rpcbTable)
colnames(rpcbXtable) <- c("(P, X, E)",
                          "$\\that_o$ [95\\% CI]",
                          "$\\that_r$ [95\\% CI]",
                          "$\\that_m$ [95\\% CI]",
                          "$c$", #"$\\sigma^2_o/\\sigma^2_r$",
                          "$d$", #"$\\that_r/\\that_o$",
                          "$p_o$",
                          "$p_r$",
                          "$p_m$",
                          "$p_Q$",
                          "$p_{\\text{S}}$"
                          )
colnames(rpcbXtable) <- paste0("\\multicolumn{1}{c}{", colnames(rpcbXtable), "}")
print(rpcbXtable, booktabs = TRUE, floating = FALSE, include.rownames = TRUE,
      sanitize.text.function = function(x){x},
      size = "small")
@
%
}
\end{table}

@@

To illustrate reverse-Bayes inference, let us return to the replication setting.
Assume we want to conduct inference about the unknown effect size $\theta$ based
on the effect estimate from the original study $\that_{o}$ and its standard
error $\sigma_{o}$. We will assume that $\that_{o}$ is normally distributed
around the unknown effect size $\theta$ with (known) variance equal to its
squared standard error $\sigma^{2}_{o}$, here and henceforth denoted by
$\that_{o} \given \theta \sim \Nor(\theta, \sigma^{2}_{o})$. Furthermore, we
specify a zero-mean normal prior with variance $\tau^{2}$ for the effect size
$\theta$, representing the position of a sceptic who does not believe in the
presence of a genuine effect. The ``stubbornness'' of the sceptic is determined
by how small the variance $\tau^{2}$ is chosen (the smaller $\tau^{2}$, the more
scepticism). Combining the sceptical prior with the likelihood produces a
posterior which is again normal
$\theta \given \that_{o}, \sigma_{o} \sim \Nor(\mu_{\text{post}}, \sigma^{2}_{\text{post}})$
with mean and variance
\begin{align*}
  &\mu_{\text{post}} = \frac{\that_{o}}{1 + \sigma^{2}_{o}/\tau^{2}}&
  &\text{and}&
  &\sigma^{2}_{\text{post}} = \frac{1}{1/\sigma^{2}_{o} + 1/\tau^{2}}.&
\end{align*}
The associated $(1- \alpha)$ highest posterior density credible interval is
given by
\begin{align}
  \left[\mu_{\text{post}} \pm z_{\alpha/2} \, \sigma_{\text{post}}\right]
  \label{eq:postCI}
\end{align}
and if this credible interval excludes parameter values smaller/larger than zero
(depending on the orientation of the effect size) this may be interpreted as
evidence\footnote{For readers who do not agree with this notion of evidence, do
  not panic. In this thesis (Paper II and IV) we extend this reverse-Bayes
  procedure to use alternative measures of evidence, such as Bayes factors.} for
a genuine effect found in the original study.

@@

This thesis consists of six papers divided into three parts. The first part
(Paper I, II, and III) focuses on methodology for design and analysis of
replication studies. The second part (Paper IV and V) revolves around
reverse-Bayes methodology with a broader scope of applications than replication
studies. % Paper IV is a review paper about
% reverse-Bayes methodology and Paper V is a short comment on another article
% which proposed a reverse-Bayes method.
The last part (Paper VI) is about integrity issues in methodological research
with simulations studies. % The article lists and illustrates questionable
% research practices in simulation studies, and gives recommendations how to avoid
% them.

>>>>

% How can we know if a finding from a scientific study is really true? For
% example, how can we know if the protective effect of a vaccine found in a study
% is real? The answer to such questions is of high importance to scientists,
% decision makers, and the general population. However, because any study result
% comes with uncertainty we can never know for sure.
How can we know whether we can trust a finding from a scientific study? For
example, how can we know whether the protective effect of a vaccine found in a
study is a genuine effect or just a spurious finding? The answer to such
questions is of high importance to scientists, decision makers, and the general
population, and one way to obtain it is through \emph{replication studies}.

A replication study consists of repeating the original study as close as
possible but with new subjects. Such a study may yield similar results which
would make us more confident about the original finding, or may yield
conflicting results which would lower our confidence. As such, replication
studies are an essential part of the scientific method as they provide a means
for substantiating genuine research findings and refuting research findings
which occurred merely by chance. For this reason, ``successful'' replication is
often a requirement, for instance, for acceptance of newly proposed scientific
theories (\eg{} a new physical model) or the implementation of policies based on
scientific knowledge (\eg{} market approval of a drug). The results from
replication studies may thus have real world consequences, % , such as deciding
% whether we receive the recommendation to get vaccinated (a consequence that
% anyone who experienced the COVID-19 pandemic is very aware of)
and it is therefore important that methodological aspects of replication
studies, such as their statistical design and analysis, are well understood.

@@

In the midst of this crisis, % , particularly from the hard-hit
% field of psychology,
various reforms were implemented to prevent an escalation of the situation % and
% save the reputation of science
\citep[for an overview see \eg{}][]{Munafo2017}. For instance, some journals
adopted the ``registered report'' format \citep{Chambers2021} in which a study
proposal is peer reviewed \emph{before} the study is conducted, and which, if
accepted, gives provisional publication acceptance regardless of the study
outcome. Similarly, digital infrastructure platforms, such as Zenodo
(\url{https://zenodo.org}) or the Open Science Framework (\url{https://osf.io}),
were created to facilitate preregistration, preprints, code and data sharing,
all of which have substantially increased over the last decade
\citep{Kidwell2016, Nosek2018, Rawlinson2019}. The practice of conducting
replication studies has also gained popularity. Several academic journals and
funders are now explicitly promoting replication research \citep{NWO2016,
  NSL2018, Nature2022}. Figure~\ref{fig:scholar} (left) illustrates this trend
by the yearly number of results for the search term ``replication study''
obtained from the search engine Web of Science. The numbers have been rapidly
growing, especially since the mid-2000s. This trend could possibly be explained
by the fact that more research results are published each year but also when the
numbers are normalized by the number of results for the search term ``study'',
the increasing trend is still visible.

@@

\begin{table}[!htb]
  \centering
  \caption{Study-level summary statistics for an original and a replication
    study. The cumulative distribution function of the standard normal
    distribution is denoted by $\Phi(\cdot)$, and the $1 - \alpha$ quantile of
    the standard normal distribution is denoted by
    $\Phi^{-1}(1 - \alpha) = \z{\alpha}$. Confidence intervals and
    \textit{p}-values are based on the assumption that the effect estimate
    $\that$ is normally distributed around the unknown effect size $\theta$ with
    (known) variance equal to the squared standard error $\sigma^{2}$.}
  \label{tab:meta}
  \begin{tabular}{lcc}
    \toprule
    & Original study & Replication study \\
    \midrule
    effect estimate & $\that_o$ & $\that_{r}$ \\
    standard error & $\sigma_{o}$ & $\sigma_{r}$ \\
    $(1 - \alpha)$ confidence interval &
    % $[\that_{o} \pm \z{\alpha/2} \sigma_{o}]$ &
    % $[\that_{r} \pm \z{\alpha/2} \sigma_{r}]$ \\
    $[\that_{o} - \z{\alpha/2} \sigma_{o}, \that_{o} + \z{\alpha/2} \sigma_{o}]$ &
    $[\that_{r} - \z{\alpha/2} \sigma_{r}, \that_{r} + \z{\alpha/2} \sigma_{r}]$ \\
    $z$-value & $z_{o} = \that_{o}/\sigma_{o}$ & $z_{r} = \that_{r}/\sigma_{r}$ \\
    $p$-value (two-sided) & $p_{o} = 2\{1 - \Phi(|z_{o}|)\}$ & $p_{r} = 2\{1 - \Phi(|z_{r}|)\}$ \\
    % One-sided $p$-value & $p_{o} = 1 - \Phi(|z_{o}|)$ & $p_{r} = 2\{1 - \Phi(|z_{r}|)\}$ \\
    \bottomrule
  \end{tabular}
\end{table}

Most methods for analyzing replication studies can be formulated in terms of
study-level summary statistics as shown in Table~\ref{tab:meta}. % Effect
% estimates, standard error, confidence intervals, and $p$-values
All of these are routinely reported in research articles, and if one of them is
missing they typically can be calculated from the others. Using summary
statistics is also often the only possible way for conducting the replication
success analysis as the raw data from the original study may not be available to
the researchers conducting the replication study. The most important statistic
is the effect estimate $\that$. It provides an estimate of the underlying effect
size $\theta$ which quantifies the true effect or association of an
intervention/exposure with an outcome variable. Typical effect sizes are mean
differences and correlations (for continuous outcomes), odds ratios (for binary
outcomes), or hazard ratios (for time-to-event outcomes). Depending on the type
of effect size, a transformation might be required so that the assumption of
approximately normally distributed effect estimates around the unknown effect
size $\theta$ (for large enough sample sizes) is justified. This could be, for
instance, the Fisher $z$-transformation for correlations or the
log-transformation for odds/hazard ratios \citep[chapter 11]{Cooper2019}. The
associated standard error $\sigma$ represents the statistical uncertainty of the
estimate. Typically, the available standard error $\sigma$ is only an estimate
of the true standard error. However, in most parts of this thesis it will be
assumed that $\sigma$ corresponds to the true standard error. This is the same
assumption as in ordinary fixed-effect/random-effects meta-analysis, and
typically reasonable if the sample size of the studies is not too small. Under
the assumption of (asymptotic) normality, confidence intervals for $\theta$ and
$p$-values for testing the null hypothesis $H_{0} \colon \theta = 0$ can be
computed as shown in Table~\ref{tab:meta}.

@@

Table~\ref{tab:rsdef} lists commonly used criteria for replication success in
terms of the summary statistics from Table~\ref{tab:meta}. The most popular
criterion is listed first and defines replication success by simultaneous
statistical significance of original and replication study along with their
effect estimates showing the same direction. This approach is also called the
``two-trials rule'' \citep{Senn2008} in drug regulation or ``vote-counting'' in
meta-analysis \citep{Cooper2019}. The criterion can similarly be defined through
one-sided $p$-values, so that the original and replication effect estimate are
not explicitly required to show the same direction as this is already taken into
account by the one-sided $p$-values. Some replication projects
\citep{Opensc2015, Errington2021} also defined replication success via
simultaneous non-significance of original and replication study
($p_{o} > \alpha$ and $p_{r} > \alpha$). However, with this definition
``replication success'' can almost always be achieved by conducting original and
replication study with very few samples so that the $p$-values are large. The
approach is also logically questionable as it could be seen as an instance of
the ``absence of evidence fallacy'' \citep{Altman1995} meaning the failure to
find evidence against the null hypothesis is erroneously interpreted as evidence
for the null hypothesis. Meta-analytic extensions of the significance approach
define replication success by significance of a combined effect
estimate.
% \footnote{The meta-analytic approach can also be given a Bayesian
% interpretation: A flat prior distribution for the underlying effect size
% $\theta$ is updated by the data from original and replication study.
% Replication success defined by a meta-analytic $p$-value being smaller than
% $\alpha$ is then equivalent to replication success via a Bayesian posterior
% probability $\Pr(\theta > 0 \given \that_{o}, \that_{r}) > 1 - \alpha/2$,
% respectively, $\Pr(\theta < 0 \given \that_{o}, \that_{r}) > 1 - \alpha/2$,
% depending on the direction of the combined estimate.}
Typically, the assumption of a common underlying effect size is seen as
reasonable so that fixed-effect meta-analysis is used for pooling.
Random-effects meta-analysis has primarily been used if more than one
replication study of the same original study are conducted since replicators are
often interested in understanding between-replication heterogeneity \citep[\eg{}
in][]{Klein2018}.

@@

\begin{figure}[!htb]
<< "RPCB-graph", fig.height = 3.5 >>=
za <- qnorm(p = 0.975)
rpcb <- read.csv("data/rpcb.csv") |>
    mutate(id = paste0("(", paper, ", ", experiment, ", ", effect, ")"),
           zo = smdo/so,
           zr = smdr/sr,
           c = so^2/sr^2,
           d = smdr/smdo,
           po2 = 2*(1 - pnorm(q = abs(zo))),
           pr2 = 2*(1 - pnorm(q = abs(zr))),
           sm = 1/sqrt(1/so^2 + 1/sr^2),
           smdm = (smdo/so^2 + smdr/sr^2)*sm^2,
           pm2 = 2*(1 - pnorm(q = abs(smdm/sm))),
           Q = (smdo - smdr)^2/(so^2 + sr^2),
           pQ = pchisq(q = Q, df = 1, lower.tail = FALSE),
           rsig = factor(pr2 <= 0.05, levels = c(FALSE, TRUE),
                         labels = c("italic(p[r]) > 0.05",
                                    "italic(p[r]) <= 0.05")))
## compute sceptical p-value (TODO fix problems with NA)
ind <- !is.na(rpcb$zo) & !is.na(rpcb$zr) & !is.na(rpcb$c)
rpcb$ps2 <- NA
rpcb$ps2[ind] <- pSceptical(zo = rpcb$zo[ind], zr = rpcb$zr[ind], c = rpcb$c[ind],
                            alternative = "two.sided", type = "nominal")

@@

upset(rpcbUpset, sets = c("Significance", "Meta analytic \nsignificance",
                          "Relative \neffect size", "Prediction \ninterval"),
      order.by = "freq", mainbar.y.label = "Successful replications (combined criteria)",
      sets.x.label = "Successful replications \n(single criterion)", #, empty.intersections = "on")
      queries = list(list(query = intersects,
                          params = c("Significance",
                                     "Meta analytic \nsignificance",
                                     "Relative \neffect size",
                                     "Prediction \ninterval"),
                          color = "#1b9e77", active = TRUE)))
@
\caption{UpSet plot for data on 132 effects from the Reproducibility Project:
  Cancer Biology \citep{Errington2021} for which effect estimates and standard
  errors are available on the standardized mean difference scale. Shown are the
  number of replication successes according to the different criteria from
  Table~\ref{tab:rsdef} and their combinations. A threshold of
  $\alpha = \Sexpr{round(100*alpha, 2)}\%$ and a relative effect size threshold
  of $d_{\text{min}} = \Sexpr{dmin}$ are used for replication success.}
\label{fig:upset}
\end{figure}

@@

Taken together, this analysis demonstrates that conclusions based on commonly
used replication success criteria often differ. It is not always clear-cut
whether or not a replication is successful. Reducing replicability to a single
criterion without mentioning these difficulties, as often done by the popular
press \citep[\eg{} ``more than half of the findings did not hold up when
retested'' in][]{Carey2015}, is a simplification of the matter.

\begin{landscape}
\begin{table}
  \centering
  \caption{Subset of results from the Reproducibility Project: Cancer Biology
    \citep{Errington2021}. (P, X, E) denotes effect number E from experiment
    number X, from original paper number P. Shown are original $\that_{o}$,
    replication $\that_{r}$, and pooled effect estimate $\that_{m}$ with 95\%
    confidence intervals, variance ratio $c = \sigma^{2}_{o}/\sigma^{2}_{r}$,
    relative effect estimate $d = \that_{r}/\that_{o}$, original
    \textit{p}-value $p_{o}$, replication \textit{p}-value $p_{r}$,
    meta-analytic \textit{p}-value $p_{m}$, $Q$-test \textit{p}-value $p_{Q}$,
    and sceptical \textit{p}-value $p_{\mathrm{S}}$. A replication success
    (green) threshold of $\alpha = 5\%$ is used for the \textit{p}-values and
    $d_{\text{min}} = 1$ for the relative effect estimate.}
  \label{tab:RPCB}
\resizebox{1.02\linewidth}{!}{%
<< "RPCB-table2", results = "asis" >>=
## function to format CI
ciFormat. <- function(lower, upper, dollar = FALSE, digits = 2) {
    if (dollar) x <- "$" else x <- ""
    paste0("[", x, format(round(lower, digits), nsmall = digits), x,
           " to ", x, format(round(upper, digits), nsmall = digits), x, "]")
}
ciFormat <- Vectorize(FUN = ciFormat.)

## function to format p-value
pFormat. <- function(p, digits = 2, dollar = FALSE, phantom = FALSE) {
    if (is.na(p)) {
        out <- NA
    } else {
        if (p > 0.0001) {
            if (p < 10^(-digits+1)) {
                out <- as.character(signif(p, digits = digits))
            } else {
                out <- format(round(p, digits), nsmall = digits)
            }
            if (phantom) out <- paste0("\\phantom{<~}", out)
        } else {
            out <- "< 0.0001"
        }
        if (dollar) {
            out <- paste0("$", out, "$")
        }
    }
    return(out)
}
pFormat <- Vectorize(FUN = pFormat.)

## randomly select a subset with one effect per paper
set.seed(7)
rpcbTable <- rpcb |>
    na.omit() |>
    group_by(paper) |>
    sample_n(size = 1) |>
    ungroup() |>
    arrange(pr2) |>
    mutate(smdoFormat = signif(smdo, 3),
           smdrFormat = signif(smdr, 3),
           smdmFormat = signif(smdm, 3),
           dFormat = tableFormat(x = d, outFun = dFormat,
                                 threshold = 1, col1 = "dark2orange",
                                 col2 = "dark2green"),
           cioL = smdo - za*so,
           cioU = smdo + za*so,
           cirL = smdr - za*sr,
           cirU = smdr + za*sr,
           cimL = smdm - za*sm,
           cimU = smdm + za*sm,,
           po2Format = tableFormat(x = po2),
           pr2Format = tableFormat(x = pr2),
           pm2Format = tableFormat(x = pm2),
           ps2Format = tableFormat(x = ps2),
           pQFormat = tableFormat(x = pQ),
           cioFormat = ciFormat(lower = cioL, upper = cioU, dollar = TRUE),
           cirFormat = ciFormat(lower = cirL, upper = cirU, dollar = TRUE),
           cimFormat = ciFormat(lower = cimL, upper = cimU), dollar = TRUE) |>
    select(id,
           smdoFormat, cioFormat,
           smdrFormat, cirFormat,
           smdmFormat, cimFormat,
           c, dFormat,
           po2Format, pr2Format, pm2Format, pQFormat, ps2Format
           )
rpcbXtable <- xtable(rpcbTable)
colnames(rpcbXtable) <- c("(P, X, E)",
                          "$\\that_o$", "[95\\% CI]",
                          "$\\that_r$", "[95\\% CI]",
                          "$\\that_m$", "[95\\% CI]",
                          "$c$",
                          "$d$",
                          "$p_o$",
                          "$p_r$",
                          "$p_m$",
                          "$p_Q$",
                          "$p_{\\text{S}}$"
                          )

colnames(rpcbXtable) <- paste0("\\multicolumn{1}{c}{", colnames(rpcbXtable), "}")
print(rpcbXtable, booktabs = TRUE, floating = FALSE, include.rownames = TRUE,
      sanitize.text.function = function(x){x},
      math.style.negative = getOption("xtable.math.style.negative", TRUE),
      size = "small")
@
%
}
\end{table}

@@

To illustrate reverse-Bayes inference, let us return to the replication setting.
Assume we want to conduct inference about the unknown effect size $\theta$ based
on the effect estimate from the original study $\that_{o}$ and its standard
error $\sigma_{o}$. We will assume that $\that_{o}$ is normally distributed
around the unknown effect size $\theta$ with (known) variance equal to its
squared standard error $\sigma^{2}_{o}$, here and henceforth denoted by
$\that_{o} \given \theta \sim \Nor(\theta, \sigma^{2}_{o})$. Furthermore, we
specify a zero-mean normal prior with variance $\tau^{2}$ for the effect size
$\theta$, representing the position of a sceptic who does not believe in the
presence of a genuine effect. The ``stubbornness'' of the sceptic is determined
by how small the variance $\tau^{2}$ is chosen (the smaller $\tau^{2}$, the more
scepticism). Combining the sceptical prior with the likelihood produces a
posterior which is again normal
$\theta \given \that_{o}, \sigma_{o} \sim \Nor(\mu_{\text{post}}, \sigma^{2}_{\text{post}})$
with mean and variance
\begin{align*}
  &\mu_{\text{post}} = \frac{\that_{o}}{1 + \sigma^{2}_{o}/\tau^{2}}&
  &\text{and}&
  &\sigma^{2}_{\text{post}} = \frac{1}{1/\sigma^{2}_{o} + 1/\tau^{2}}.&
\end{align*}
The associated $(1- \alpha)$ highest posterior density credible interval has
limits
\begin{align}
  \left[\mu_{\text{post}} - z_{\alpha/2} \, \sigma_{\text{post}},
  \mu_{\text{post}} + z_{\alpha/2} \, \sigma_{\text{post}}\right]
  \label{eq:postCI}
\end{align}
and if this credible interval excludes parameter values smaller/larger than zero
(depending on the orientation of the effect size) this may be interpreted as
evidence for a genuine effect found in the original study. There exist also
other definitions of statistical evidence and in latter parts of this thesis
(Paper\ref{PaperII} and Paper\ref{PaperIV}) we will extend this reverse-Bayes
procedure to employ alternative measures of evidence, such as Bayes factors.

@@

This thesis consists of six papers divided into three parts. The first part
(Paper~\ref{PaperI}, \ref{PaperII}, and \ref{PaperIII}) focuses on methodology
for design and analysis of replication studies. The second part
(Paper~\ref{PaperIV} and \ref{PaperV}) revolves around reverse-Bayes methodology
with a broader scope of applications than replication
studies. % Paper IV is a review paper about
% reverse-Bayes methodology and Paper V is a short comment on another article
% which proposed a reverse-Bayes method.
The last part (Paper~\ref{PaperVI}) is about integrity issues in methodological
research with simulations
studies. % The article lists and illustrates questionable
% research practices in simulation studies, and gives recommendations how to avoid
% them.
